---
output:
  pdf_document: default
  html_document: default
---
---
title: "Customer Retention Intelligence - Statistical Analysis"
author: "Abhinav Sriram"
output: html_document
---


# Overview

This analysis evaluated relationships identified during exploratory data 
analysis and determines which customer characteristics remain associatd with 
churn after accounting for other factors.

This analysis focuses on: 
-Contract type
-Customer tenure
-Monthly charges
-Premium technical support
-Internet type
-Customer referrals
-Demographic characteristics

```{r setup, message=FALSE, warning=FALSE}
library(DBI)
library(RSQLite)
library(dplyr)
library(broom)
library(here)


knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE
)

```

#Load Data from SQLite

The analytical data is retrieved directly from the SQLite database created
during the data-engineering phase.

```{r}
db_path = here(
  "database", "customer_retention.db"
)

con <- dbConnect(SQLite(), db_path)
```


```{r, query-data}
query <- "
SELECT
    c.customer_id,
    c.age,
    c.married,
    c.number_of_dependents,
    c.number_of_referrals,
    c.tenure_months,
    c.churn,

    a.contract,
    a.payment_method,
    a.monthly_charge,

    s.internet_service,
    s.internet_type,
    s.premium_tech_support

FROM customers AS c

JOIN customer_accounts AS A
  ON c.customer_id = a.customer_id

JOIN customer_services AS s
  ON c.customer_id = s.customer_id

WHERE c.churn IS NOT NULL;"

customer_data <- dbGetQuery(con, query)

dbDisconnect(con)
```

#Initial Inspection

```{r inspect-data}
class(customer_data)
dim(customer_data)
head(customer_data)

```
The analytical dataset should contain 6,589 customers representing customers
who either stayed or churned. Newly joined customers are excluded because their
long-term churn outcome has not yet been observed.

#Data Preparation

Categorical variables are converted to factors before statistical testing.

```{r factors}
customer_data <- customer_data %>%
  mutate(churn = factor(churn, levels = c(0,1)), 
         contract = factor(contract), 
         premium_tech_support = factor(premium_tech_support), 
         internet_type = factor(internet_type), 
         internet_service = factor(internet_service), 
         married = factor(married), 
         payment_method = factor(payment_method)
  )
```

#Contract Type and Churn

EDA suggested that month-to-month customers experience substantially higher 
churn than customers with longer-term contracts.

The hypotheses are: 
**H0:** Contract type and churn are independent.
**HA:** Contract type and churn are associated.

```{r contract-table}
contract_table <- table(
  customer_data$contract, 
  customer_data$churn
)

contract_table
```

```{r contract-chi-square}
contract_test <- chisq.test(contract_table)

contract_test
```
#Premium Technical Support and Churn

EDA showed substantially higher churn among customers without premium technical
support.

The hypotheses are: 
**H0:** Premium technical support and churn are independent.
**HA:** Premium technical support and churn are associated.

```{r support-table}
support_table <- table(customer_data$premium_tech_support, customer_data$churn)

support_table
```
```{r support-chi-square}
support_test <- chisq.test(support_table)

support_test
```

#Internet Type and Churn

EDA showed differences in churn rates among fiber optic, cable, DSL, and
customers without internet service.

The hypotheses are: 
**H0:** Internet type and churn are independent.
**HA:** Internet type and churn are associated.

```{r internet-table}
internet_table <- table(
  customer_data$internet_type, 
  customer_data$churn
)

internet_table
```

```{r internet-chi-square}
internet_test <- chisq.test(internet_table)
internet_test
```
#Prepare Data for Logistic Regression
The invalid negative monthly charges identified during data cleaning were 
converted to missing values.

For this statistical model, observations with missing monthly charges are 
excluded

```{r model-data}
sum(is.na(customer_data$monthly_charge))

model_df <- customer_data %>%
  filter(!is.na(monthly_charge))

dim(model_df)
```

Create a numeric binary churn outcome for logistic regression.

```{r numeric-target}
model_df$churn_numeric <- as.integer(
  as.character(model_df$churn)
)
```

#Multivariable Logistic Regression

A logistic regression model is used to evaluate which variables remain
associated with churn after accounting for other predictors.

```{r logistic-model}
churn_model <- glm(churn_numeric ~
                     contract + 
                     tenure_months +
                     monthly_charge + 
                     premium_tech_support + 
                     internet_type + 
                     number_of_referrals +
                     age + 
                     number_of_dependents +
                     married, 
                   data = model_df, 
                   family = binomial)

summary(churn_model)
```

#Odds Ratios

Logistic regression coefficients are exponentiated to produce odds ratios.

An odds ratio: 
-Greater than 1 indicate higher churn odds.
-Less than 1 indicated lower churn odds.
-Near 1 indicates little change in churn odds.

```{r odds-ratios}
odds_ratios <- tidy(churn_model, exponentiate = TRUE, coef.int = TRUE)

odds_ratios
```


#Statistically Significant Predictors

The following table displays predictors with p-values below 0.05.

```{r}
significant_predictors <- odds_ratios %>%
  filter(
    term != "(Intercept)", 
    p.value < 0.05
  ) %>%
  arrange(p.value)

significant_predictors
```

#Save Results

The logistic regression results are exported for use in later project stages 
and documentation.

```{r save-results}
write.csv(
  odds_ratios, 
  here(
    "data", "processed", "logistic_refression_results.csv"
  ), 
  row.names = FALSE
)
```













